Self Regulation Ontology Retest Data: Initial Exploration

### 2017-02-21 SRO_Retest_Analyses.Rmd
cont working on retest exploration

retest_data <- retest_data %>%
  group_by(sub_id) %>%
  do(correct_subids(.,test_worker_lookup))

#Are all retest subjects in test_data? No. 
sum(retest_data$sub_id %in% test_data$sub_id) == nrow(retest_data)

#Who is missing?
retest_data$sub_id[which(retest_data$sub_id %in% test_data$sub_id == FALSE)]
```

@@

```{r}
retest_data <- as.data.frame(retest_data)
retest_subs_test_data <- as.data.frame(retest_subs_test_data)
matching_dv_columns <- c()
for(i in 1:length(matching_columns)){
  if(is.numeric(retest_data[,matching_columns[i]]) & is.numeric(retest_subs_test_data[,matching_columns[i]])){
    matching_dv_columns <- c(matching_dv_columns, matching_columns[i])
  }
}

@@

Distribution of Pearson correlations
```{r}
ggplot(cor_df, aes(pearson))+
  geom_histogram()+
  theme_bw()
```

Distribution of spearman correlations
```{r}
ggplot(cor_df, aes(spearman))+
  geom_histogram()+
  theme_bw()
```

@@

Distributions of correlations by task vs. survey
```{r}
cor_df %>%
  gather(key, value, -dv, -task) %>%
  ggplot(aes(value))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(key~task)
```

@@

Mean reliability for surveys
```{r}
cor_df %>%
  filter(task == 'task') %>%
  select(dv, spearman, pearson) %>%
  arrange(-pearson, -spearman) %>%
  summarise(mean_pearson = mean(pearson),
            mean_spearman = mean(spearman))
```

@@

Mean reliability for surveys
```{r}
cor_df %>%
  filter(task == 'survey') %>%
  select(dv, spearman, pearson) %>%
  arrange(-pearson, -spearman) %>%
  summarise(mean_pearson = mean(pearson),
            mean_spearman = mean(spearman))
```

@@

```{r}
library(psych)
corMat <- cor(rest_subs_test_mngf_z)
solution <- fa(r = corMat, nfactors = 2, rotate = "oblimin", fm = "pa")
```

Mean reliabilities for different kinds of variables (e.g. drift rates and NDs; separately for EZ and HDDM; subtraction vs basic variables - single vs. multiple)
```{r}
cor_df$dv
```

>>>>

retest_data <- retest_data %>%
  group_by(sub_id) %>%
  do(correct_subids(.,test_worker_lookup))
```

Are all retest subjects in test_data? No. 
```{r}
sum(retest_data$sub_id %in% test_data$sub_id) == nrow(retest_data)
```

Who is missing?
```{r}
retest_data$sub_id[which(retest_data$sub_id %in% test_data$sub_id == FALSE)]
```

@@

```{r}
retest_data <- as.data.frame(retest_data)
retest_subs_test_data <- as.data.frame(retest_subs_test_data)

matching_dv_columns <- c()

for(i in 1:length(matching_columns)){
  if(is.numeric(retest_data[,matching_columns[i]]) & is.numeric(retest_subs_test_data[,matching_columns[i]])){
    matching_dv_columns <- c(matching_dv_columns, matching_columns[i])
  }
}

@@

Distribution of all Pearson and Spearman correlations
```{r}
cor_df %>%
  gather(key, value) %>%
  ggplot(aes(value))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(~key)
```

@@

Distributions of correlations by task vs. survey

Survey reliabilities are noticably higher than task reliabilities.

Task reliability distributions looks alsmost bimodal. **Should figure out whether the lower ones are specific tasks or if they are certain kinds of vars (e.g. certain ddm parameters)**
```{r}
cor_df %>%
  gather(key, value, -dv, -task) %>%
  ggplot(aes(value))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(key~task)
```

@@

Mean reliability for tasks
```{r}
cor_df %>%
  filter(task == 'task') %>%
  select(dv, spearman, pearson) %>%
  arrange(-pearson, -spearman) %>%
  summarise(mean_pearson = mean(pearson),
            mean_spearman = mean(spearman),
            num_vars = n())
```

@@

Mean reliability for surveys
```{r}
cor_df %>%
  filter(task == 'survey') %>%
  select(dv, spearman, pearson) %>%
  arrange(-pearson, -spearman) %>%
  summarise(mean_pearson = mean(pearson),
            mean_spearman = mean(spearman),
            num_vars = n())
```

Check if the variables that went in to the correlation matrix have normal distributions since at least the Pearson correlations depend on this assumption. If not transform data columns in both retest and test data and calculate a new correlation df.

Function to check if the variables that went in to cor_df have normal distributions
```{r}
diff_from_normal <- function(col){
  return(shapiro.test(col)$p.value < 0.05)
}
```

Check distributions of retest data

The smaller meaningful vars have been somewhat cleaned. So far in this notebook we looked at all the vars in their rawest form.

Create and organize new df where distribution of vars data is stored. First for retest data only; then for test data for subjects who have retest data - restricting here since these are the subsets of data that went in to the correlation matrix. Merge and cleanup
```{r}
dist_retest_data <- as.data.frame(apply(retest_data[,matching_dv_columns], 2, diff_from_normal))
names(dist_retest_data) <- c('retest_not_normal')
dist_retest_data$dv <- row.names(dist_retest_data)
row.names(dist_retest_data) <- seq(1:nrow(dist_retest_data))
dist_retest_data = dist_retest_data %>% select(dv, retest_not_normal)

dist_retest_subs_test_data <- as.data.frame(apply(retest_subs_test_data[,matching_dv_columns], 2, diff_from_normal))
names(dist_retest_subs_test_data) <- c('test_not_normal')
dist_retest_subs_test_data$dv <- row.names(dist_retest_subs_test_data)
row.names(dist_retest_subs_test_data) <- seq(1:nrow(dist_retest_subs_test_data))
dist_retest_subs_test_data = dist_retest_subs_test_data %>% select(dv, test_not_normal)

dist_data <- merge(dist_retest_data, dist_retest_subs_test_data, by = 'dv')

rm(dist_retest_data, dist_retest_subs_test_data)
```

Add whether dv's are tasks or survey to distribution dfs
```{r}
dist_data = cor_df %>%
  select(dv, task) %>%
  left_join(dist_data, by = 'dv')
```

What proportion is not normal?
A large proportion of the variables are not normal for data from both sources. Larger proportions of survey data are not normal compared to task data. 
```{r}
dist_data %>%
  group_by(task) %>%
  summarise(mean_retest_not_normal  = mean(retest_not_normal),
            mean_test_not_normal = mean(test_not_normal))
```

Are the same vars that are not normally distributed in the retest data not normal in test data as well?
List of variables that have different distributions in the two time points (shown here is whether they are normally distributed in the retest data)
A majority of them are drift diffusion parameters (but note that these vars are overrepresented to begin with - more on them later).
```{r}
dist_data %>%
  filter(retest_not_normal != test_not_normal)
```

Apply the same cleaning procedure that was used for meaningful vars for all vars:
```{r}
transform_remove_skew = function(data, threshold = 1){
}
```

Transform non normal vars
Transform both retest and test data if not normally distributed in either. 
Order of transformations tried: log and sqrt
```{r}
#Add columns to dist_data that will be updated
dist_data = dist_data %>%
  mutate(retest_transform = ifelse(retest_not_normal, NA, 'none'),
         test_transform = ifelse(test_not_normal, NA, 'none'))

get_transform <- function(data){
  t_data = data
  transform = 'none'
  if(diff_from_normal(data) & !is.na(diff_from_normal(data))){
    t_data = log1p(data)
    transform = 'log1p'
    if(diff_from_normal(t_data) & !is.na(diff_from_normal(t_data))){
      t_data = sqrt(data)
      transform = 'sqrt'
      if(diff_from_normal(t_data) & !is.na(diff_from_normal(t_data))){
        transform = 'failed'
      }
    }
  }
  return(transform)
}

for(i in 1:nrow(dist_data)){
  if(is.na(dist_data$retest_transform[i])){
    dist_data$retest_transform[i] = get_transform(retest_data[,dist_data$dv[i]])
  }
  if(is.na(dist_data$test_transform[i])){
    dist_data$test_transform[i] = get_transform(retest_subs_test_data[,dist_data$dv[i]])
  }
}

t_retest_data = retest_data
t_retest_subs_test_data = retest_subs_test_data

#check if normal
#if yes stop and record transformation = none
#if transform: log
#check if normal
#if yes stop and record transformation = log
#if no transform: sqrt
```

Mean reliabilities for different kinds of variables (e.g. drift rates and NDs; separately for EZ and HDDM; subtraction vs basic variables - single vs. multiple)
```{r}
```

--------------------------------

Can you get the factor structure for the retest
step 1:Ttest on all measures of T1 for people w and without retest
```{r}

```

step2 : efa on the people you don't have retest on and then predict how well this holds up for people w retest on t1 and on t2

@@

```{r}
corMat <- rest_subs_test_mngf_z %>% select(-sub_id) %>% cor(.,use = 'complete.obs')

solution <- fa(r = corMat, nfactors = 3, rotate = "oblimin", fm = "pa")
```

```{r}
solution
```
